Geometrization of characters of tori over non-Archimedean local fields

### 2013-08-09 shanghai.tex
Improving the proof of the proposition on finite character sheaves.

%\begin{proposition}\label{tensor}
The category $\CS(G)$ of character sheaves on $G$ is a rigid monoidal
category [insert reference] under
$\cs{L} \otimes \cs{L'} \ceq (\gcs{L}\otimes\gcs{L'}, \mu\otimes\mu', F\otimes F')$
with duals given by applying the functor $\sheafHom(\ - \ ,(\EE)_{\bG})$ (sheaf hom) to
conditions \ref{CS.0}, \ref{CS.1} and \ref{CS.2}.
%\end{proposition}
(The category of characters sheaves on $G$ is not abelian, so $\CS(G)$
is not a tensor category in the sense of \cite{Deligne:tensorielles}.)
%\begin{proof}\end{proof}
%
This rigid monoidal category structure for $\CS(G)$ puts a group
structure on the set $\CSiso{G}$ of isomorphism classes in
$\CS(G)$; we refer to this as the \cdef{Picard group of character
sheaves} on $G$; {\it cf.} {\tt http://ncatlab.org/nlab/show/Picard+group}.
\Clifton{Replace ncatlab with other references.}

@@

\begin{proof}
  Let $\cs{L}$ be a character sheaf on $G$. We start by showing that
  $\bf^*\gcs{L}$ is locally constant on connected components of
  $\bH$. Let $c_j :\bH^j \hookrightarrow \bH$ be any
  connected component and let $i$ be the image of $j$ under the group
  homomorphism $\pi_0(\bG) \to \pi_0(\bH)$ obtained by
  applying the functor $\pi_0$ to $\bf$; see \cite{vdG&M}*{III, 3.28}, for example.
  \Clifton{Replace with a better reference, eventually.}
  Write $c_i : \bG^i \hookrightarrow \bG$ for
  the inclusion of $\bG^i$ and $\bf^j : \bH^j \to \bG^i$
  for the restriction of $f$ to $\bH^j$.  Since $\cs{L}$
  is a character sheaf, $\gcs{L}$ is locally constant on the connect
  components of $\bG$; thus, there is a finite etale covering
  $a_i : X_i \to \bG^i$ such that $a_i^* (\gcs{L}\vert_{\bG^i})$
  is constant. Let $b_j : Y_j \to \bH^j$ be the
  pull-back of $a_i$ along $\bf^j$; let $g_j : Y_j \to X_i$ be
  the map opposite $\bf^j$. Then $b_j$ is a finite etale covering
  of $\bH^j$ and
  \[
  b_j^* \left( (\bf^* \gcs{L})\vert_{\bH^j}\right)
  = (\bf\circ c_j \circ b_j)^*\gcs{L}
  = (c_i\circ a_i\circ g_j)^*\gcs{L} = g_j^* a_i^* (\gcs{L}\vert_{\bG^i}).
  \]

@@

\begin{proof}
\todo{This proof is missing.}
\end{proof}

@@

Let $\CSb(G)$ be the category of pairs $(\cs{L}_0,\mu_0)$ 
where $\cs{L}_0$ is a constructible $\ell$-adic sheaf on $G$, 
which is a rank-$1$ local system on each connected component of $G$, 
equipped with an isomorphism $\mu_0 : m^* \cs{L}_0 \to \cs{L}_0 \boxtimes \cs{L}_0$ 
satisfying the analogue of Condition~\ref{CS.1} on $G$; 
morphisms in $\CSf(G)$ are defined as in (the second part of) Condition~\ref{CS.2}. 
This is the category of \cdef{bounded character sheaves} on $G$. 
This is a rigid monoidal category in the obvious way, 
with duals given by $\cs{L}_0^\vee \ceq \sheafHom(\cs{L}_0,(\EE)_{G})$.

@@

To define the category of finite character sheaves we must work a bit harder.
Consider the following category $C(G)$:
an object in $C(G)$ is a pair $(f,\psi)$ where: 
$f : H\to G$ is an isogeny (a finite, surjective, etale, morphism of commutative group $\Fq$-schemes) 
for which the action of $\Gal(\bFq/\Fq)$ on the etale group scheme $\ker f$ is trivial, 
and $\psi : \ker f\to \Aut(V)$ is a representation on a finite-dimensional $\EE$-vector space $V$;
a map $(f,\psi) \to (f',\psi')$ in $C(G)$ is a pair $(\varphi,T)$ 
where $\varphi$ is a morphism of schemes such that $f' = f\circ \varphi$
and $T : V\to V'$ is a linear transformation equivariant 
for the action of $\ker f'$ on $V'$ by $\psi'$ 
and the action of $\ker f'$ on $V$ by $\psi \circ \varphi$.
%
We say that $(\varphi,T)$ in $C(G)$ is a weak isomorphism if $T$ is an isomorphism of vector spaces. 
Let $C(G)[W^{-1}]$ be the localization of $C(G)$ at weak isomorphisms.
Now, let $C_1(G)$ be the subcategory of $C(G)$ such that $\psi$ is a one-dimensional representation
and let $\CSf(G)$ be the localisation of $C_1(G)$ at weak isomorphisms. 
This is the category of \cdef{finite character sheaves} on $G$.

\Clifton{Add the monoidal category structure here.}

@@

\begin{proof}
 To prove the proposition we exhibit a full and faithful monoidal functor $L_G : \CSf(G) \to \CSb(G)$.
 %
 We will do this by constructing a faithful monoidal functor $L : C_1(G) \to \CSb(G)$ 
 such that ${\bar L}(\varphi,T)$ is an isomorphism precisely when $(\varphi,T)$ is a weak isomorphism. 
 Factoring along $C_1(G) \to \CSb(G)$ will produce the full and faithful monoidal functor we seek.

 We start by putting a tensor category structure on $C(G)$. 
 Let $(f,\psi)$ and $(f',\psi')$ be objects in $C(G)$. 
 The category of isogenies to $G$ admits products: 
 write $f\times_G f' : H\times_G H' \to G$ for the product of $f : H\to G$ and $f' : H'\to G$. 
 Since $\ker(f\times_G f') = (\ker f)\times (\ker f')$, 
 the tensor product of the representations $\psi : \ker f\to \Aut(V)$ and $\psi' : \ker f'\to \Aut(V')$ 
 is the representation $\psi \otimes\psi' :  \ker(f\times_G f') \to \Aut(V\otimes V')$. 
 This allows us to set $(f,\psi)\otimes(f',\psi') \ceq (f\times_G f' ,\psi\otimes\psi')$. 
 The same fact makes sense of $(f,\psi)\oplus(f',\psi')\ceq (f\times_G f' ,\psi\oplus\psi')$. 
 The dual of $(f,\psi)$ is $(f,\psi)^\vee \ceq (f,\psi^\vee)$ 
 where $\psi^\vee : \ker f \to \Aut(V^\vee)$ and $V^\vee = \Hom_\text{vec}(V,\EE)$. 
 The verification that these operations make $C(G)$ a tensor category is routine.

 Now we introduce a faithful functor $L : C(G) \to \Loc(G)$ to the category of local systems on $G$.
 For any object $(f,\psi)$ in $C(G)$, let $L(f,\psi)$ be the $\psi$-isotypic component of $f_* V_H$, 
 where $V_H$ refers to $V$ as a constant sheaf on $H$. 
 Note that we may identify the stalk $(f_* V_H)_{\bar g}$ of $f_* V_H$ at a geometric point ${\bar g}$ on $G$ 
 with the $\EE$-vector space $\Hom_\text{set}(\abs{f^{-1}({\bar g})},V)$ 
 and the stalk $L(f,\psi)_{\bar g}$ of $L(f,\psi)$ at a geometric point ${\bar g}$ on $G$ 
 with the $\EE$-vector space consisting of those functions $s: \abs{f^{-1}({\bar g})} \to V$ such that 
 $s(a\cdot {\bar h}) = \psi(a)(s({\bar h}))$ for all $a\in \ker f$ and for all ${\bar h} \in f^{-1}({\bar g})$.
 Since $L(f,\psi)$ is trivialized by $f$, it is a local system on $G$.
 For any map $(\varphi,T)$ in $C(G)$, let $L(\varphi,T) : L(f,\psi) \to L(f',\psi')$ be the homomorphism of local systems defined 
 %at geometric points $L(\varphi,T)_{\bar g} : L(f,\psi)_{\bar g} \to L(f',\psi')_{\bar g}$ 
 by $s \mapsto s' \ceq T\circ s\circ \varphi$. 
 This defines a monoidal functor $L$ from $C(G)$ to $\Loc(G)$. 
 This functor commutes with pull-back along any morphism of group schemes.

>>>>

%\begin{proposition}\label{tensor}
The category $\CS(G)$ of character sheaves on $G$ is a rigid monoidal
category [insert reference] under
$\cs{L} \otimes \cs{L'} \ceq (\gcs{L}\otimes\gcs{L'}, \mu\otimes\mu', F\otimes F')$
with duals given by applying the functor 
\Clifton{Should this be $\sheafHom(\ - \ ,\EE)$?} $\sheafHom(\ - \ ,(\EE)_{\bG})$ (sheaf hom) to
conditions \ref{CS.0}, \ref{CS.1} and \ref{CS.2}.
%\end{proposition}
(The category of characters sheaves on $G$ is not abelian, so $\CS(G)$
is not a tensor category in the sense of \cite{Deligne:tensorielles}.)
%\begin{proof}\end{proof}
%
This rigid monoidal category structure for $\CS(G)$ puts a group
structure on the set $\CSiso{G}$ of isomorphism classes in
$\CS(G)$; we refer to this as the \cdef{Picard group of character
sheaves} on $G$; {\it cf.} {\tt http://ncatlab.org/nlab/show/Picard+group}.
\Clifton{Replace ncatlab with other references.}

@@

\begin{proof}
  Let $\cs{L}$ be a character sheaf on $G$. We start by showing that
  $\bf^*\gcs{L}$ is locally constant on connected components of
  $\bH$. Let $c_j :\bH^j \hookrightarrow \bH$ be any
  connected component and let $i$ be the image of $j$ under the group
  homomorphism $\pi_0(\bG) \to \pi_0(\bH)$ obtained by
  applying the functor $\pi_0$ to $\bf$; see \cite{vdG&M}*{III, 3.28}, for example.
  \Clifton{I haven't found a better reference for this yet!}
  Write $c_i : \bG^i \hookrightarrow \bG$ for
  the inclusion of $\bG^i$ and $\bf^j : \bH^j \to \bG^i$
  for the restriction of $f$ to $\bH^j$.  Since $\cs{L}$
  is a character sheaf, $\gcs{L}$ is locally constant on the connect
  components of $\bG$; thus, there is a finite etale covering
  $a_i : X_i \to \bG^i$ such that $a_i^* (\gcs{L}\vert_{\bG^i})$
  is constant. Let $b_j : Y_j \to \bH^j$ be the
  pull-back of $a_i$ along $\bf^j$; let $g_j : Y_j \to X_i$ be
  the map opposite $\bf^j$. Then $b_j$ is a finite etale covering
  of $\bH^j$ and
  \[
  b_j^* \left( (\bf^* \gcs{L})\vert_{\bH^j}\right)
  = (\bf\circ c_j \circ b_j)^*\gcs{L}
  = (c_i\circ a_i\circ g_j)^*\gcs{L} = g_j^* a_i^* (\gcs{L}\vert_{\bG^i}).
  \]

@@

\begin{proof}
\Clifton{This proof is missing! Can David insert it?}
\end{proof}

@@

Let $\CSb(G)$ be the category of pairs $(\cs{L}_0,\mu_0)$ 
where $\cs{L}_0$ is a constructible $\ell$-adic sheaf on $G$, 
which is a rank-$1$ local system on each connected component of $G$, 
equipped with an isomorphism $\mu_0 : m^* \cs{L}_0 \to \cs{L}_0 \boxtimes \cs{L}_0$ 
satisfying the analogue of Condition~\ref{CS.1} on $G$; 
morphisms in $\CSf(G)$ are defined as in (the second part of) Condition~\ref{CS.2}. 
This is the category of \cdef{bounded character sheaves} on $G$. 
\Clifton{Should this be $\sheafHom(\cs{L}_0,\EE)$?}
This is a rigid monoidal category in the obvious way, 
with duals given by $\cs{L}_0^\vee \ceq \sheafHom(\cs{L}_0,(\EE)_{G})$.

@@

To define the category of finite character sheaves we must work a bit harder.
Consider the following category $C(G)$:
an object in $C(G)$ is a pair $(f,\psi)$ where: 
$f : H\to G$ is an isogeny (a finite, surjective, etale, morphism of commutative group $\Fq$-schemes) 
for which the action of $\Gal(\bFq/\Fq)$ on the etale group scheme $\ker f$ is trivial, 
and $\psi : \ker f\to \Aut(V)$ is a representation on a finite-dimensional $\EE$-vector space $V$;
a map $(f,\psi) \to (f',\psi')$ in $C(G)$ is a pair $(\varphi,T)$ 
where $\varphi$ is a morphism of schemes such that $f' = f\circ \varphi$
and $T : V\to V'$ is a linear transformation equivariant 
for the action of $\ker f'$ on $V'$ by $\psi'$ 
and the action of $\ker f'$ on $V$ by $\psi \circ \varphi$.

We put a tensor category structure on $C(G)$ as follows.
 Let $(f,\psi)$ and $(f',\psi')$ be objects in $C(G)$. 
 The category of isogenies to $G$ admits products: 
 write $f\times_G f' : H\times_G H' \to G$ for the product of $f : H\to G$ and $f' : H'\to G$. 
 Since $\ker(f\times_G f') = (\ker f)\times (\ker f')$, 
 the tensor product of the representations $\psi : \ker f\to \Aut(V)$ and $\psi' : \ker f'\to \Aut(V')$ 
 is the representation $\psi \otimes\psi' :  \ker(f\times_G f') \to \Aut(V\otimes V')$. 
 This allows us to set $(f,\psi)\otimes(f',\psi') \ceq (f\times_G f' ,\psi\otimes\psi')$. 
 The same fact makes sense of $(f,\psi)\oplus(f',\psi')\ceq (f\times_G f' ,\psi\oplus\psi')$. 
 The dual of $(f,\psi)$ is $(f,\psi)^\vee \ceq (f,\psi^\vee)$ 
 where $\psi^\vee : \ker f \to \Aut(V^\vee)$ and $V^\vee = \Hom_\text{vec}(V,\EE)$. 
 The verification that these operations make $C(G)$ a tensor category is routine.

We say that $(\varphi,T)$ in $C(G)$ is a weak isomorphism if $T$ is an isomorphism of vector spaces. 
Let $C(G)[W^{-1}]$ be the localization of $C(G)$ at weak isomorphisms.
Now, let $C_1(G)$ be the subcategory of $C(G)$ such that $\psi$ is a one-dimensional representation
and let $\CSf(G)$ be the localisation of $C_1(G)$ at weak isomorphisms. 
This is the rigid symmetric monoidal category of \cdef{finite character sheaves} on $G$.

@@

\begin{proof}
 To prove the proposition we exhibit a full and faithful monoidal functor $L_G : \CSf(G) \to \CSb(G)$.
 %
 We will do this by constructing a faithful monoidal functor $L : C_1(G) \to \CSb(G)$ 
 such that $L(\varphi,T)$ is an isomorphism precisely when $(\varphi,T)$ is a weak isomorphism. 
 Factoring along $C_1(G) \to \CSb(G)$ will produce the full and faithful monoidal functor we seek.

 Now we introduce a faithful functor $L : C(G) \to \Loc(G)$ to the category of local systems on $G$.
 For any object $(f,\psi)$ in $C(G)$, let $L(f,\psi)$ be the $\psi$-isotypic component of $f_* V_H$, 
 where $V_H$ refers to $V$ as a constant sheaf on $H$. 
 We may identify the stalk $(f_* V_H)_{\bar g}$ of $f_* V_H$ at a geometric point ${\bar g}$ on $G$ 
 with the $\EE$-vector space $\Hom_\text{set}(\abs{f^{-1}({\bar g})},V)$ 
 and the stalk $L(f,\psi)_{\bar g}$ of $L(f,\psi)$ at a geometric point ${\bar g}$ on $G$ 
 with the $\EE$-vector space consisting of those functions $s: \abs{f^{-1}({\bar g})} \to V$ such that 
 $s(a\cdot {\bar h}) = \psi(a)(s({\bar h}))$ for all $a\in \ker f$ and for all ${\bar h} \in f^{-1}({\bar g})$.
 Since $L(f,\psi)$ is trivialized by $f$, it is a local system on $G$.
 For any map $(\varphi,T)$ in $C(G)$, let $L(\varphi,T) : L(f,\psi) \to L(f',\psi')$ be the homomorphism of local systems defined 
 %at geometric points $L(\varphi,T)_{\bar g} : L(f,\psi)_{\bar g} \to L(f',\psi')_{\bar g}$ 
 by $s \mapsto s' \ceq T\circ s\circ \varphi$. 
 This defines a monoidal functor $L$ from $C(G)$ to $\Loc(G)$. 
 This functor commutes with pull-back along any morphism of group schemes.
